---
title: 'Reproducible Research Week 2 Project'
author: 'Marcel M.'
date: 'Feb 22nd, 2020'
output:
  html_document:
    keep_md: true
---


# Reproducible Research - Week 2 Project

============================================================================================

## 1) Code for reading in the data set
```{r}
setwd("C:/Users/Marcel/Documents/DataScience/Course5")
if(!file.exists("Project1")) {
  
   dir.create("Project1")}

setwd("Project1")
unzip("repdata_data_activity.zip")

library(readr)

repdata <- read_csv("activity.csv")
```


## 2) Histogram of the total number of steps taken each day.
```{r}
total_steps <- aggregate(repdata$steps, by= list(repdata$date), FUN=sum, na.rm= TRUE)
colnames(total_steps) <- c("Date", "Steps")

library(ggplot2)


ggplot(total_steps, aes(Steps)) + geom_histogram(color="orange", fill="yellow", binwidth = 2000) + labs(x="Total Steps", y="Frequency") + ggtitle("Total Steps Taken Each Day")
```


## 3) Mean and median number of steps taken each day.
```{r}
meanSteps <- mean(total_steps$Steps, na.rm = TRUE)
print(meanSteps)

medianSteps <- median(total_steps$Steps, na.rm = TRUE)
print(medianSteps)
```


## 4) Time series plot of the average number of steps taken.
```{r}
library(dplyr)

timeSeries <- repdata %>% group_by(date) %>% summarize(Average.Steps = mean(steps, na.rm = TRUE))


ggplot(timeSeries, aes(x= date, y= Average.Steps)) + 
      geom_line(size=1, color="darkblue") +
      labs(x = "Date", y ="Average Steps") +
      ggtitle("Average Steps per Day")
```


## 5) The 5-minute interval that, on average, contains the maximum number of steps.
```{r}
repdata %>% group_by(interval) %>% summarize(Average.Steps = mean(steps, na.rm = TRUE)) %>% filter(Average.Steps == max(Average.Steps))
```


## 6) Code to describe and show a strategy for imputing missing data.
```{r}
totalNA <- sum(is.na(repdata))
totalNA

intervalMean <- repdata %>% group_by(interval) %>% summarize(Mean.Interval = mean(steps,  na.rm = TRUE))

repdata$complete.steps <- ifelse(is.na(repdata$steps),              round(intervalMean$Mean.Interval[match(repdata$interval,   intervalMean$interval)],0),repdata$steps)

repdataComplete <- data.frame(repdata$complete.steps, repdata$date, repdata$interval)
colnames(repdataComplete) <- c("Steps", "Date", "Interval")
head(repdataComplete)
```


## 7) Histogram of the total number of steps taken each day after missing values are imputed.
```{r}
stepsDay <- repdataComplete %>% group_by(Date) %>% summarize(Total.Steps = sum(Steps))


ggplot(stepsDay, aes(Total.Steps)) + geom_histogram(color="red", fill="magenta", binwidth = 2000) + labs(x= "Total Steps", y="Frequency") + ggtitle("Total Steps Taken Each Day")


meanStepsComplete <- mean(stepsDay$Total.Steps)
print(meanStepsComplete)

medianStepsComplete <- median(stepsDay$Total.Steps)
print(medianStepsComplete)
```


## 8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.
```{r}
repdataComplete$Day.Name <- weekdays(repdataComplete$Date)

repdataComplete$Classification <- ifelse(repdataComplete$Day.Name == "Saturday" | repdataComplete$Day.Name == "Sunday", "Weekend", "Weekday")
head(repdataComplete)

timeSeriesClassification <- repdataComplete %>% group_by(Interval, Classification) %>% summarize(Interval.Steps.Mean = mean(Steps))
head(timeSeriesClassification)


ggplot(timeSeriesClassification, aes(Interval, Interval.Steps.Mean)) + geom_line(size=1.5, color= "darkblue") + facet_grid(Classification~.) + labs(x="Interval", y="Average Steps") + ggtitle("Average Steps Taken Across Weekdays and Weekends")
```

